\part{Turing Machines and Decidability}

\section{Turing Machines}

\begin{enumerate}[resume]
	\item \Level{2} Write a low-level description of a TM to recognize the language:
	\[
	\{w \in \{0, 1\}^\star : \#_0(w) = (\#_1(w))^2 \}.
	\]
	
	\item \Level{2} A \emph{backtracking Turing Machine} (BTM) is just like a normal TM except that there is another transition that ``backtracks''--when taken from state $q$, this transition instantly resets the tape contents and tape head location to be exactly when the machine was last in state $q$. If the machine was not in state $q$ before, the tape does not change.
	
	Prove that the languages recognized by BTMs are equivalent to those recognized by ``standard'' (deterministic, single-tape) TMs.
	
	\item \Level{2} A \emph{stack Turing machine} (STM) is like a normal TM except in every tape cell there is, in addition to a symbol, a stack. On each transition taken, the machine, in addition to modifying the cell's symbol and move left or right, can push or pop from that individual cell's stack (or both, or neither, as how a PDA can). If the STM attempts to pop from a cell's stack that is empty, the computation terminates.
	
	Prove that the languages recognized by STMs are equivalent to those recognized by ``standard'' (deterministic, single-tape) TMs.
	
\end{enumerate}

\section{Decidable Languages}

\begin{enumerate}[resume]
	
\item \Level{2} Let $L$ be the language:
\begin{align*}
\{ \langle D, N, R, G, P \rangle :\;\text{$D$ is a DFA, $N$ is an NFA, $R$ is a regex, $G$ is a regular grammar,}\\
\text{$P$ is a PDA, and $L(P) \subseteq ((L(D) \cap L(N)) \cup \overline{L(R)}) L(G)$ }   \}
\end{align*}
Show that $L$ is decidable (Hint: what is the intersection of a CFL and a regular language?)

\item \Level{2} Let $WEIRD_{NFA} = \{ \langle M \rangle : M$ is an NFA, $M$ has at most 7 states, accepts $\varepsilon$, and $L(M)$ is infinite$\}$.
Show that $WEIRD_{NFA}$ is decidable.

\item \Level{2} Let $L$ be a language, and let $\textsf{Power}(L)$ be:
\[
\{ w^n : w \in L, n \ge 0 \}.
\]
Prove that if $L$ is recognizable, then $\textsf{Power}(L)$ is also recognizable.

\item \Level{3} I'm concerned that the NFAs we usually deal with are ``too large'' and that we can make them even simpler. I want to determine, given an NFA $N$, whether or not there is some other NFA with fewer states than that also recognizes the language $L(N)$. 

However, this may not be ``minimal'' enough, because there could be still another NFA with fewer transitions than $N$ still with the same number of states and language $L(N)$. 

Formally state this problem. Then show that this problem is decidable.

\item \label{regex_crossword} \Level{3} You are given a grid $A[1..m][1..n]$ ($m$ rows, $n$ columns) as well as $m$ regular expressions $R_1, \cdots, R_m$ corresponding to each row, and $n$ more regular expressions $C_1, \cdots, C_n$ corresponding to each column. The alphabet of all regexes is $\Sigma$. Your task is to determine if there is a way of assigning a character in $\Sigma$ in each of the cells of $A$ such that the concatenation of the cells in row 1 (from left to right) is a string in $L(R_1)$, row 2's concatenation is a string in $L(R_2)$, as well as for all the other row and column regexes (columns concatenate from top to bottom). We say that this input has a \emph{solution} if there is such an assignment. 

For example, here is a graphical representation of a possible input with $\Sigma = \{a, b\}$:

\begin{table}[!htbp]
	\centering
	\begin{tabular}{llll}
		\cline{2-4}
		\multicolumn{1}{l|}{$(bba)^\star$} & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} \\ \cline{2-4} 
		\multicolumn{1}{l|}{$b(a^\star b)^\star$} & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} \\ \cline{2-4} 
		&     $(a \cup bb)^\star$                 &       $(ba)^\star ab$                &   $((ba)^\star \cup (ab)^\star)^\star$                
	\end{tabular}
\end{table}

Let the language $\textsf{RegexCrossword}$ be the set of all instances $\langle A, R_1, \cdots, R_m, C_1, \cdots, C_n\rangle$ where $A$ is an $m \times n$ grid, $R_1, \cdots, R_m, C_1, \cdots, C_n$ are regexes, and there is a solution to $A$ as described above. Show that $\textsf{RegexCrossword}$ is decidable. (Hint: brute-force.)

\item \Level{4} Let $\textsf{RegexCrosswordWord}$ be the language similar to that in \cref{regex_crossword}, except now the cells can be filled with a string in $\Sigma^\star$ instead of just a single character. Is $\textsf{RegexCrosswordWord}$ decidable? Recognizable? 

\end{enumerate}